---
title: "Reproducible Research Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Loading and preprocessing the data
Let's load the packages we'll be using for this project:

```{r load-packages, message=FALSE}
library(dplyr)
library(ggplot2)
```

In the chunk of code below we've:

        1. Read the data set and kept it in the `activity`variable.  
        2. Used the `as.Date()` function to convert the character vector `date`to date type.  
        3. Used the `as.factor` function to have the weekdays as levels.  
        4. Added a new variable and called it `weekday` containing the days according to the date.  

Lately, we've summarized all the variables to have a global view of our dataset.

```{r}

activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)

str(activity)

weekd <- as.factor(weekdays(activity$date))

activity <- activity %>%
        mutate(weekday = weekd)

summary(activity)

```
The variables included in this dataset are:  

              steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)  
              date: The date on which the measurement was taken in YYYY-MM-DD format.  
              interval: Identifier for the 5-minute interval in which measurement was taken.  

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

## Plots

### What is mean total number of steps taken per day?

```{r}
steps_per_day <- aggregate(steps ~ date, activity, sum, na.rm = TRUE)
ggplot(steps_per_day, aes(x = steps)) + 
        geom_histogram(boundary=0, binwidth=2500, col="darkblue",
                       fill="lightblue") +
        ggtitle("Histogram of steps per day")

head(steps_per_day)
steps_per_day %>%
        summarise(mean_steps = mean(steps, na.rm = TRUE), med_steps = median(steps, na.rm = TRUE))


```

### What is the average daily activity pattern?

```{r}
steps_per_time <- aggregate(steps ~ interval, activity, mean, na.action=na.omit)

steps_per_time$time <- steps_per_time$interval/100

ggplot(steps_per_time, aes(time, steps)) + 
        geom_line(col="darkgreen")+
        ggtitle("Average steps per time interval")+
        xlab("Time")+
        ylab("Steps")+
        theme(plot.title = element_text(size = 12))


```

